---
title: "CompareID"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CompareID}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(compareid)
```



# Basic example

We're going to play with the package data, `table1` and `table2`.

```{r results = 'asis', echo = F}
knitr::kable(table1)
```

```{r results = 'asis', echo = F}
knitr::kable(table2)
```


## Data preparation

### Creating the object

We have to select the column that we want to work with. 
The default join type is left.

```{r}
ids <- CompareID$new(table1, table2, "municipi")

print(ids, show_fields = TRUE)
```

## Steps

If we print the object right away, we can see the miss matches.

This shows the fields from the first table, that where not found at the second table.

```{r}
# ids$join_miss
ids
```

To find the fields that didn't match from the second table:

```{r}
ids$join_search
```

We can see that we could apply some transformations.

- All tables: To_lower all the letters.
- 1st table: Remove ", l'" pattern
- 2nd table: Remove "l' " patter
- 2nd table: Change "alella" -> "alélla"
- 2nd table: Change "arenys mar" -> "arenys de mar"

```{r}
ids$add_step(type = "lower", which = "all")
ids$add_step(type = "replace", pattern = ", l'",
             position = "end", which = "id1")
ids$add_step(type = "replace", pattern = "l'",
             position = "start", which = "id2")
ids$add_step(type = "manual", match = c("alélla" = "alella",
                                        "arenys mar" = "arenys de mar"), which = "id1")
```

## Apply the steps

Printing the object, the steps apply automatically:

```{r}
ids
```

To explicitly apply the steps, use `join()`:

```{r}
ids$join()
ids$join_miss   # Veure els que no s'han trobat
ids$join_search # Veure els que no s'han trobat de l'altra taula
```

Using join_miss or join_search they apply too.

## Final merge:

To get the final ids result:

```{r}
ids$join_result
```

To get the complete merged data:

```{r}
ids$join_data(table1, table2)
```



# Complex example


We're going to play with the package data, `table_complex1` and `table_complex2`.

```{r results = 'asis', echo = F}
dplyr::tibble(table_complex1)
```

```{r results = 'asis', echo = F}
dplyr::tibble(table_complex2)
```


## Data preparation

### Creating the object

Load the data to the object:

```{r}
dades <- CompareID$new(table_complex1,
                       table_complex2,
                       "municipi")
```

```{r}
dades
```

We have 143 missing matches!

If we `View(dades$join_miss)`, we can see the miss matches in detail.

### Apply steps

#### The first try

One strategy is having a `View(dades$join_search)` window to search a pattern, and 
filter something similar to the first miss.

In our case, if we filter by "Alam" we find:

```{r}
dades$join_search |> 
  dplyr::filter(stringr::str_detect(id, "Alam"))
```

We see a pattern, where some prepositions are at the end. We could delete
this prepositions and see if we improve our matches a bit:

```{r}
dades$add_step(type = "replace",
               which = "id2",
               pattern = c(", el", ", la", ", els", ", les"),
               position = "end")

dades
```

There are only 35 miss matches now!


#### Modify a step

We can se that there is an extra preposition ", l'", and ", Es".

We can create the object again, or delete the step and add a new one.

```{r}
dades$steps_id2$remove_step(1)
dades$add_step(type = "replace",
               which = "id2",
               pattern = c(", el", ", la", ", els", ", les", ", l'", ", Es"),
               position = "end")

dades
```


#### Analisis of miss matches one by one

A lot of the remaining municipalities are not from catalonia (Alfajarín, Biescas...) so we could remove them at the start, when we construct the object.

For specific matches, we could do a manual step:

```{r}
dades$add_step(type = "manual",
               which = "id2",
               match = c("Brunyola i Sant Martí Sapresa" = "Brunyola",
                         "Calonge i Sant Antoni" = "Calonge"))

dades
```

The miss matches are reduced to 16!

If there are a lot more miss matches and don't see any more patterns,we could stop and try to use the package [fuzzyjoin](https://cran.r-project.org/web/packages/fuzzyjoin/readme/README.html) to get the miss match distance, and analize / filter from there.
